---
title: "Compare-Alpha diversity-HumGut"
output: html_notebook
---
#Part 2 Alpha diversity （HumGut data）

This notebook creats for HumGut
But also includes typical data for comparison

```{r message=FALSE, warning=FALSE, include=FALSE}
library(phyloseq)
library(tidyverse)
library(dplyr)
library(microbiome)
library(hrbrthemes)
library(knitr)
library(viridis)
library(fantaxtic)
```
Two phyloseq projects from HumGut:
PhySeq_KrakenHG_06.09.21.rds
PhySeq_BrakenHG_06.09.21.rds
```{r}
#typical
PhySeq_Kraken_20.07.21 <- readRDS("Kraken/PhySeq_Kraken_20.07.21.rds")
PhySeq_Braken_20.07.21 <- readRDS("Kraken-Braken/PhySeq_Braken_20.07.21.rds")
PhySeq_Metaphlan_20.07.21 <- readRDS("Metaphlan3/PhySeq_Metaphlan_20.07.21.rds")

#HG
PhySeq_KrakenHG_06.09.21 <- readRDS("Kraken-HumGut/PhySeq_KrakenHG_06.09.21.rds")
PhySeq_BrakenHG_06.09.21 <- readRDS("Braken-HumGut/PhySeq_BrakenHG_06.09.21.rds")
```

##Alpha diversity visualization
Prepare
```{r}
####################################typical############################################
#kraken
aggre_kraken_phy <- aggregate_taxa(PhySeq_Kraken_20.07.21,"species") %>% aggregate_rare(level = "species", detection = 0.1/100, prevalence = 1/100)
infant_aggre_kraken_phy <- subset_samples(aggre_kraken_phy, !(sample_type %in% c("F","M"))) %>% microbiome::transform(transform = "compositional")
#kraken_allknown
aggre_kraken_allknown_phy <- aggregate_taxa(PhySeq_Kraken_20.07.21,"species") %>% aggregate_rare(level = "species", detection = 0.1/100, prevalence = 1/100) %>% subset_taxa(!(species=="Unknown"))
infant_aggre_kraken_allknown_phy <- subset_samples(aggre_kraken_allknown_phy, !(sample_type %in% c("F","M"))) %>% microbiome::transform(transform = "compositional")
#Braken
aggre_braken_phy <- aggregate_taxa(PhySeq_Braken_20.07.21,"species") %>% aggregate_rare(level = "species", detection = 0.1/100, prevalence = 1/100)
infant_aggre_braken_phy <- subset_samples(aggre_braken_phy, !(sample_type %in% c("F","M"))) %>% microbiome::transform(transform = "compositional")
#Metaphlan
aggre_meta_phy <- aggregate_taxa(PhySeq_Metaphlan_20.07.21,"species") %>% aggregate_rare(level = "species", detection = 0.1/100, prevalence = 1/100)
infant_aggre_meta_phy <- subset_samples(aggre_meta_phy, !(sample_type %in% c("F","M"))) %>% microbiome::transform(transform = "compositional")

#################################################HG###############################
#kraken_HG
aggre_krakenHG_phy <- aggregate_taxa(PhySeq_KrakenHG_06.09.21,"species") %>% aggregate_rare(level = "species", detection = 0.1/100, prevalence = 1/100)
infant_aggre_krakenHG_phy <- subset_samples(aggre_krakenHG_phy, !(sample_type %in% c("F","M"))) %>% microbiome::transform(transform = "compositional")

#kraken_HG_allknown
aggre_krakenHG_allknown_phy <- aggregate_taxa(PhySeq_KrakenHG_06.09.21,"species") %>% aggregate_rare(level = "species", detection = 0.1/100, prevalence = 1/100) %>% subset_taxa(!(species=="Unknown"))
infant_aggre_krakenHG_allknown_phy <- subset_samples(aggre_krakenHG_allknown_phy, !(sample_type %in% c("F","M"))) %>% microbiome::transform(transform = "compositional")

#Braken_HG
aggre_brakenHG_phy <- aggregate_taxa(PhySeq_BrakenHG_06.09.21,"species") %>% aggregate_rare(level = "species", detection = 0.1/100, prevalence = 1/100)
infant_aggre_brakenHG_phy <- subset_samples(aggre_brakenHG_phy, !(sample_type %in% c("F","M"))) %>% microbiome::transform(transform = "compositional")

```

Alpha diversity
```{r}
################################typical####################################
#kraken
alpha__kraken <-microbiome::alpha(infant_aggre_kraken_phy, index = c("observed","shannon","chao1"))
kable(head(alpha__kraken))
#kraken_allknown
alpha__kraken_allknown <-microbiome::alpha(infant_aggre_kraken_allknown_phy, index = c("observed","shannon","chao1"))
kable(head(alpha__kraken_allknown))
#braken
alpha__braken <-microbiome::alpha(infant_aggre_braken_phy, index = c("observed","shannon","chao1"))
kable(head(alpha__braken))
#meta
alpha__meta <-microbiome::alpha(infant_aggre_meta_phy, index = c("observed","shannon","chao1"))
kable(head(alpha__meta))

#####################################HG########################################
#krakenHG
alpha__krakenHG <-microbiome::alpha(infant_aggre_krakenHG_phy, index = c("observed","shannon","chao1"))
kable(head(alpha__krakenHG))

#krakenHG_allknown
alpha__krakenHG_allknown <-microbiome::alpha(infant_aggre_krakenHG_allknown_phy, index = c("observed","shannon","chao1"))
kable(head(alpha__krakenHG_allknown))

#brakenHG
alpha__brakenHG <-microbiome::alpha(infant_aggre_brakenHG_phy, index = c("observed","shannon","chao1"))
kable(head(alpha__brakenHG))

```

boxplot by time flow
```{r}
###############################typical########################################
#kraken
sample_kraken <- sample_data(infant_aggre_kraken_phy)
box_kraken <- merge(sample_kraken, alpha__kraken, by = 0) %>% data.frame()

p_box_kraken <- box_kraken %>% ggplot(aes(x=sample_type, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "Kraken", caption = "shannon") +
  ylim(c(0,4)) +
  theme(plot.title = element_text(size=11))
print(p_box_kraken)

#kraken_allknown
sample_kraken_allknown <- sample_data(infant_aggre_kraken_allknown_phy)
box_kraken_allknown <- merge(sample_kraken_allknown, alpha__kraken_allknown, by = 0) %>% data.frame()

p_box_kraken_allknown <- box_kraken_allknown %>% ggplot(aes(x=sample_type, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "Kraken_allknown", caption = "shannon") +
  ylim(c(0,4)) +
  theme(plot.title = element_text(size=11))
print(p_box_kraken_allknown)

#braken
sample_braken <- sample_data(infant_aggre_braken_phy)
box_braken <- merge(sample_braken, alpha__braken, by = 0) %>% data.frame()

p_box_braken <- box_braken %>% ggplot(aes(x=sample_type, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "Braken", caption = "shannon") +
  ylim(c(0,4)) +
  theme(plot.title = element_text(size=11))
print(p_box_braken)

#meta
sample_meta <- sample_data(infant_aggre_meta_phy)
box_meta <- merge(sample_meta, alpha__meta, by = 0) %>% data.frame()

p_box_meta <- box_meta %>% ggplot(aes(x=sample_type, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "Meta", caption = "shannon") +
  ylim(c(0,4)) +
  theme(plot.title = element_text(size=11))
print(p_box_meta)

###########################HG################################################
#krakenHG
sample_krakenHG <- sample_data(infant_aggre_krakenHG_phy)
box_krakenHG <- merge(sample_krakenHG, alpha__krakenHG, by = 0) %>% data.frame()

p_box_krakenHG <- box_krakenHG %>% ggplot(aes(x=sample_type, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "Kraken_HG", caption = "shannon") +
  ylim(c(0,4)) +
  theme(plot.title = element_text(size=11))
print(p_box_krakenHG)

#krakenHG_allknown
sample_krakenHG_allknown <- sample_data(infant_aggre_krakenHG_allknown_phy)
box_krakenHG_allknown <- merge(sample_krakenHG_allknown, alpha__krakenHG_allknown, by = 0) %>% data.frame()

p_box_krakenHG_allknown <- box_krakenHG_allknown %>% ggplot(aes(x=sample_type, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "Kraken_HG_allknown", caption = "shannon") +
  ylim(c(0,4)) +
  theme(plot.title = element_text(size=11))
print(p_box_krakenHG_allknown)

#brakenHG
sample_brakenHG <- sample_data(infant_aggre_brakenHG_phy)
box_brakenHG <- merge(sample_brakenHG, alpha__brakenHG, by = 0) %>% data.frame()

p_box_brakenHG <- box_brakenHG %>% ggplot(aes(x=sample_type, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "Braken_HG", caption = "shannon") +
  ylim(c(0,4)) +
  theme(plot.title = element_text(size=11))
print(p_box_brakenHG)
```

boxplot compare just among HumGut
```{r}
#B1
box_krakenHG_B1 <- filter(box_krakenHG, sample_type == "B1") %>% select(diversity_shannon)
box_krakenHG_B1$name <- "KrakenHG"
box_krakenHG_allknown_B1 <- filter(box_krakenHG_allknown, sample_type == "B1") %>% select(diversity_shannon)
box_krakenHG_allknown_B1$name <- "KrakenHG_allknown"
box_brakenHG_B1 <- filter(box_brakenHG, sample_type == "B1") %>% select(diversity_shannon)
box_brakenHG_B1$name <- "BrakenHG"

box_B1 <- rbind(box_krakenHG_B1, box_krakenHG_allknown_B1) %>% rbind(box_brakenHG_B1)

p_box_B1 <- box_B1 %>% ggplot(aes(x=name, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "HumGut", caption = "shannon") +
  xlab("B1")
  theme(plot.title = element_text(size=11))
print(p_box_B1)

#B4
box_krakenHG_B4 <- filter(box_krakenHG, sample_type == "B4") %>% select(diversity_shannon)
box_krakenHG_B4$name <- "KrakenHG"
box_krakenHG_allknown_B4 <- filter(box_krakenHG_allknown, sample_type == "B4") %>% select(diversity_shannon)
box_krakenHG_allknown_B4$name <- "KrakenHG_allknown"
box_brakenHG_B4 <- filter(box_brakenHG, sample_type == "B4") %>% select(diversity_shannon)
box_brakenHG_B4$name <- "BrakenHG"

box_B4 <- rbind(box_krakenHG_B4, box_krakenHG_allknown_B4) %>% rbind(box_brakenHG_B4)

p_box_B4 <- box_B4 %>% ggplot(aes(x=name, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "HumGut", caption = "shannon") +
  xlab("B4")
  theme(plot.title = element_text(size=11))
print(p_box_B4)

#B5
box_krakenHG_B5 <- filter(box_krakenHG, sample_type == "B5") %>% select(diversity_shannon)
box_krakenHG_B5$name <- "KrakenHG"
box_krakenHG_allknown_B5 <- filter(box_krakenHG_allknown, sample_type == "B5") %>% select(diversity_shannon)
box_krakenHG_allknown_B5$name <- "KrakenHG_allknown"
box_brakenHG_B5 <- filter(box_brakenHG, sample_type == "B5") %>% select(diversity_shannon)
box_brakenHG_B5$name <- "BrakenHG"

box_B5 <- rbind(box_krakenHG_B5, box_krakenHG_allknown_B5) %>% rbind(box_brakenHG_B5)

p_box_B5 <- box_B5 %>% ggplot(aes(x=name, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "HumGut", caption = "shannon") +
  xlab("B5")
  theme(plot.title = element_text(size=11))
print(p_box_B5)

#B7
box_krakenHG_B7 <- filter(box_krakenHG, sample_type == "B7") %>% select(diversity_shannon)
box_krakenHG_B7$name <- "KrakenHG"
box_krakenHG_allknown_B7 <- filter(box_krakenHG_allknown, sample_type == "B7") %>% select(diversity_shannon)
box_krakenHG_allknown_B7$name <- "KrakenHG_allknown"
box_brakenHG_B7 <- filter(box_brakenHG, sample_type == "B7") %>% select(diversity_shannon)
box_brakenHG_B7$name <- "BrakenHG"

box_B7 <- rbind(box_krakenHG_B7, box_krakenHG_allknown_B7) %>% rbind(box_brakenHG_B7)

p_box_B7 <- box_B7 %>% ggplot(aes(x=name, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "HumGut", caption = "shannon") +
  xlab("B7")
  theme(plot.title = element_text(size=11))
print(p_box_B7)
```

Boxplot compare among all data
```{r}
############################B1########################################
#typical
box_kraken_B1 <- filter(box_kraken, sample_type == "B1") %>% select(diversity_shannon)
box_kraken_B1$name <- "Kraken"
box_kraken_allknown_B1 <- filter(box_kraken_allknown, sample_type == "B1") %>% select(diversity_shannon)
box_kraken_allknown_B1$name <- "Kraken_allknown"
box_braken_B1 <- filter(box_braken, sample_type == "B1") %>% select(diversity_shannon)
box_braken_B1$name <- "Braken"
#HG
box_krakenHG_B1 <- filter(box_krakenHG, sample_type == "B1") %>% select(diversity_shannon)
box_krakenHG_B1$name <- "KrakenHG"
box_krakenHG_allknown_B1 <- filter(box_krakenHG_allknown, sample_type == "B1") %>% select(diversity_shannon)
box_krakenHG_allknown_B1$name <- "KrakenHG_allknown"
box_brakenHG_B1 <- filter(box_brakenHG, sample_type == "B1") %>% select(diversity_shannon)
box_brakenHG_B1$name <- "BrakenHG"
#meta
box_meta_B1 <- filter(box_meta, sample_type == "B1") %>% select(diversity_shannon)
box_meta_B1$name <- "Meta"

#plot
box_B1_all <- rbind(box_krakenHG_B1, box_krakenHG_allknown_B1, box_brakenHG_B1,
                    box_kraken_B1, box_kraken_allknown_B1, box_braken_B1,
                    box_meta_B1)

box_B1_all$name_class <- factor(box_B1_all$name, levels = c("Kraken", "Kraken_allknown", "Braken", 
                                                            "KrakenHG", "KrakenHG_allknown", "BrakenHG",
                                                            "Meta"))

p_box_B1_all <- box_B1_all %>% ggplot(aes(x=name, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "All data", caption = "shannon") +
  xlab("B1")
  theme(plot.title = element_text(size=11))
print(p_box_B1_all)

############################B4########################################
#typical
box_kraken_B4 <- filter(box_kraken, sample_type == "B4") %>% select(diversity_shannon)
box_kraken_B4$name <- "Kraken"
box_kraken_allknown_B4 <- filter(box_kraken_allknown, sample_type == "B4") %>% select(diversity_shannon)
box_kraken_allknown_B4$name <- "Kraken_allknown"
box_braken_B4 <- filter(box_braken, sample_type == "B4") %>% select(diversity_shannon)
box_braken_B4$name <- "Braken"
#HG
box_krakenHG_B4 <- filter(box_krakenHG, sample_type == "B4") %>% select(diversity_shannon)
box_krakenHG_B4$name <- "KrakenHG"
box_krakenHG_allknown_B4 <- filter(box_krakenHG_allknown, sample_type == "B4") %>% select(diversity_shannon)
box_krakenHG_allknown_B4$name <- "KrakenHG_allknown"
box_brakenHG_B4 <- filter(box_brakenHG, sample_type == "B4") %>% select(diversity_shannon)
box_brakenHG_B4$name <- "BrakenHG"
#meta
box_meta_B4 <- filter(box_meta, sample_type == "B4") %>% select(diversity_shannon)
box_meta_B4$name <- "Meta"

#plot
box_B4_all <- rbind(box_krakenHG_B4, box_krakenHG_allknown_B4, box_brakenHG_B4,
                    box_kraken_B4, box_kraken_allknown_B4, box_braken_B4,
                    box_meta_B4)

box_B4_all$name_class <- factor(box_B4_all$name, levels = c("Kraken", "Kraken_allknown", "Braken", 
                                                            "KrakenHG", "KrakenHG_allknown", "BrakenHG",
                                                            "Meta"))

p_box_B4_all <- box_B4_all %>% ggplot(aes(x=name, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "All data", caption = "shannon") +
  xlab("B4")
  theme(plot.title = element_text(size=11))
print(p_box_B4_all)

############################B5########################################
#typical
box_kraken_B5 <- filter(box_kraken, sample_type == "B5") %>% select(diversity_shannon)
box_kraken_B5$name <- "Kraken"
box_kraken_allknown_B5 <- filter(box_kraken_allknown, sample_type == "B5") %>% select(diversity_shannon)
box_kraken_allknown_B5$name <- "Kraken_allknown"
box_braken_B5 <- filter(box_braken, sample_type == "B5") %>% select(diversity_shannon)
box_braken_B5$name <- "Braken"
#HG
box_krakenHG_B5 <- filter(box_krakenHG, sample_type == "B5") %>% select(diversity_shannon)
box_krakenHG_B5$name <- "KrakenHG"
box_krakenHG_allknown_B5 <- filter(box_krakenHG_allknown, sample_type == "B5") %>% select(diversity_shannon)
box_krakenHG_allknown_B5$name <- "KrakenHG_allknown"
box_brakenHG_B5 <- filter(box_brakenHG, sample_type == "B5") %>% select(diversity_shannon)
box_brakenHG_B5$name <- "BrakenHG"
#meta
box_meta_B5 <- filter(box_meta, sample_type == "B5") %>% select(diversity_shannon)
box_meta_B5$name <- "Meta"

#plot
box_B5_all <- rbind(box_krakenHG_B5, box_krakenHG_allknown_B5, box_brakenHG_B5,
                    box_kraken_B5, box_kraken_allknown_B5, box_braken_B5,
                    box_meta_B5)

box_B5_all$name_class <- factor(box_B5_all$name, levels = c("Kraken", "Kraken_allknown", "Braken", 
                                                            "KrakenHG", "KrakenHG_allknown", "BrakenHG",
                                                            "Meta"))

p_box_B5_all <- box_B5_all %>% ggplot(aes(x=name, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "All data", caption = "shannon") +
  xlab("B5")
  theme(plot.title = element_text(size=11))
print(p_box_B5_all)

############################B7########################################
#typical
box_kraken_B7 <- filter(box_kraken, sample_type == "B7") %>% select(diversity_shannon)
box_kraken_B7$name <- "Kraken"
box_kraken_allknown_B7 <- filter(box_kraken_allknown, sample_type == "B7") %>% select(diversity_shannon)
box_kraken_allknown_B7$name <- "Kraken_allknown"
box_braken_B7 <- filter(box_braken, sample_type == "B7") %>% select(diversity_shannon)
box_braken_B7$name <- "Braken"
#HG
box_krakenHG_B7 <- filter(box_krakenHG, sample_type == "B7") %>% select(diversity_shannon)
box_krakenHG_B7$name <- "KrakenHG"
box_krakenHG_allknown_B7 <- filter(box_krakenHG_allknown, sample_type == "B7") %>% select(diversity_shannon)
box_krakenHG_allknown_B7$name <- "KrakenHG_allknown"
box_brakenHG_B7 <- filter(box_brakenHG, sample_type == "B7") %>% select(diversity_shannon)
box_brakenHG_B7$name <- "BrakenHG"
#meta
box_meta_B7 <- filter(box_meta, sample_type == "B7") %>% select(diversity_shannon)
box_meta_B7$name <- "Meta"

#plot
box_B7_all <- rbind(box_krakenHG_B7, box_krakenHG_allknown_B7, box_brakenHG_B7,
                    box_kraken_B7, box_kraken_allknown_B7, box_braken_B7,
                    box_meta_B7)

box_B7_all$name_class <- factor(box_B7_all$name, levels = c("Kraken", "Kraken_allknown", "Braken", 
                                                            "KrakenHG", "KrakenHG_allknown", "BrakenHG",
                                                            "Meta"))

p_box_B7_all <- box_B7_all %>% ggplot(aes(x=name, y=diversity_shannon)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.6, alpha=0.9)+
  labs(title = "All data", caption = "shannon") +
  xlab("B7")
  theme(plot.title = element_text(size=11))
print(p_box_B7_all)
```


##Top 10 taxa

Example:
```{r}
### get top 10 genus
physeq <- read_rds("PhySeq_BrakenHG_06.09.21.rds")
# aggregating at a taxa level and getting the compositional normalization
pseq_plot <- physeq %>%
  aggregate_taxa(level = "genus") %>%  
  microbiome::transform(transform = "compositional")

# melting the object
pmelt <- psmelt(pseq_plot)

# average Taxa by time point
av_pmelt <- pmelt %>% select(sample_type, OTU, Abundance) %>% 
  group_by(sample_type, OTU) %>% mutate(av_abundance=mean(Abundance)) %>%
  select(-Abundance) %>% unique() %>% ungroup()

# get top 10 genus by time point
top_pmelt <- av_pmelt %>% group_by(sample_type) %>% 
  slice_max(order_by = av_abundance, n = 10)
```

Prepare
note: in meta, use "genera"
```{r}
####################################typical############################################
#kraken
top_kraken_phy <- aggregate_taxa(PhySeq_Kraken_20.07.21,"genus") %>% subset_samples(!(sample_type %in% c("F","M"))) %>%
  microbiome::transform(transform = "compositional")
#kraken_allknown
top_kraken_allknown_phy <- aggregate_taxa(PhySeq_Kraken_20.07.21,"genus") %>% subset_taxa(!(genus=="Unknown")) %>% 
  subset_samples(!(sample_type %in% c("F","M"))) %>% microbiome::transform(transform = "compositional")
#Braken
top_braken_phy <- aggregate_taxa(PhySeq_Braken_20.07.21,"genus") %>% subset_samples(!(sample_type %in% c("F","M"))) %>%
  microbiome::transform(transform = "compositional")
#Metaphlan
top_meta_phy <- aggregate_taxa(PhySeq_Metaphlan_20.07.21,"genera") %>% subset_samples(!(sample_type %in% c("F","M"))) %>%
  microbiome::transform(transform = "compositional")

#################################################HG###############################
#kraken_HG
top_krakenHG_phy <- aggregate_taxa(PhySeq_KrakenHG_06.09.21,"genus") %>% subset_samples(!(sample_type %in% c("F","M"))) %>%
  microbiome::transform(transform = "compositional")
#kraken_HG_allknown
top_krakenHG_allknown_phy <- aggregate_taxa(PhySeq_KrakenHG_06.09.21,"genus") %>% subset_taxa(!(genus=="Unknown")) %>%
  subset_samples(!(sample_type %in% c("F","M"))) %>% microbiome::transform(transform = "compositional")
#Braken_HG
top_brakenHG_phy <- aggregate_taxa(PhySeq_BrakenHG_06.09.21,"genus") %>% subset_samples(!(sample_type %in% c("F","M"))) %>%
  microbiome::transform(transform = "compositional")
```

get top 10 by time point
```{r}
####################################typical############################################
#kraken
# melting the object
kraken_melt <- psmelt(top_kraken_phy)
# average Taxa by time point
av_kraken_melt <- kraken_melt %>% select(sample_type, OTU, Abundance) %>%
  group_by(sample_type, OTU) %>% mutate(av_abundance=mean(Abundance)) %>%
  select(-Abundance) %>% unique() %>% ungroup()
# get top 10 genus by time point
top_kraken <- av_kraken_melt %>% group_by(sample_type) %>%
  slice_max(order_by = av_abundance, n = 10)

#kraken_allknown
kraken_allknown_melt <- psmelt(top_kraken_allknown_phy)

av_kraken_allknown_melt <- kraken_allknown_melt %>% select(sample_type, OTU, Abundance) %>%
  group_by(sample_type, OTU) %>% mutate(av_abundance=mean(Abundance)) %>%
  select(-Abundance) %>% unique() %>% ungroup()

top_kraken_allknown <- av_kraken_allknown_melt %>% group_by(sample_type) %>%
  slice_max(order_by = av_abundance, n = 10)

#Braken
braken_melt <- psmelt(top_braken_phy)

av_braken_melt <- braken_melt %>% select(sample_type, OTU, Abundance) %>%
  group_by(sample_type, OTU) %>% mutate(av_abundance=mean(Abundance)) %>%
  select(-Abundance) %>% unique() %>% ungroup()

top_braken <- av_braken_melt %>% group_by(sample_type) %>%
  slice_max(order_by = av_abundance, n = 10)

#Metaphlan
meta_melt <- psmelt(top_meta_phy)

av_meta_melt <- meta_melt %>% select(sample_type, OTU, Abundance) %>%
  group_by(sample_type, OTU) %>% mutate(av_abundance=mean(Abundance)) %>%
  select(-Abundance) %>% unique() %>% ungroup()

top_meta <- av_meta_melt %>% group_by(sample_type) %>%
  slice_max(order_by = av_abundance, n = 10)

#################################################HG###############################
#kraken_HG
krakenHG_melt <- psmelt(top_krakenHG_phy)

av_krakenHG_melt <- krakenHG_melt %>% select(sample_type, OTU, Abundance) %>%
  group_by(sample_type, OTU) %>% mutate(av_abundance=mean(Abundance)) %>%
  select(-Abundance) %>% unique() %>% ungroup()

top_krakenHG <- av_krakenHG_melt %>% group_by(sample_type) %>%
  slice_max(order_by = av_abundance, n = 10)

#kraken_HG_allknown
krakenHG_allknown_melt <- psmelt(top_krakenHG_allknown_phy)

av_krakenHG_allknown_melt <- krakenHG_allknown_melt %>% select(sample_type, OTU, Abundance) %>%
  group_by(sample_type, OTU) %>% mutate(av_abundance=mean(Abundance)) %>%
  select(-Abundance) %>% unique() %>% ungroup()

top_krakenHG_allknown <- av_krakenHG_allknown_melt %>% group_by(sample_type) %>%
  slice_max(order_by = av_abundance, n = 10)

#Braken_HG
brakenHG_melt <- psmelt(top_brakenHG_phy)

av_brakenHG_melt <- brakenHG_melt %>% select(sample_type, OTU, Abundance) %>%
  group_by(sample_type, OTU) %>% mutate(av_abundance=mean(Abundance)) %>%
  select(-Abundance) %>% unique() %>% ungroup()

top_brakenHG <- av_brakenHG_melt %>% group_by(sample_type) %>%
  slice_max(order_by = av_abundance, n = 10)
```

visualization
```{r}
##########################################typical#############################
#kraken
p_top_kraken <- top_kraken %>% data.frame() %>% ggplot(aes(fill=OTU, y=av_abundance, x=sample_type)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(title = "Kraken",
       subtitle = "aggregate by genus",
       caption = "Top 10 taxa")

print(p_top_kraken)

#kraken_allknown
p_top_kraken_allknown <- top_kraken_allknown %>% data.frame() %>% ggplot(aes(fill=OTU, y=av_abundance, x=sample_type)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(title = "Kraken_allknown",
       subtitle = "aggregate by genus",
       caption = "Top 10 taxa")

print(p_top_kraken_allknown)

#braken
p_top_braken <- top_braken %>% data.frame() %>% ggplot(aes(fill=OTU, y=av_abundance, x=sample_type)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(title = "Braken",
       subtitle = "aggregate by genus",
       caption = "Top 10 taxa")

print(p_top_braken)

#meta
p_top_meta <- top_meta %>% data.frame() %>% ggplot(aes(fill=OTU, y=av_abundance, x=sample_type)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(title = "Meta",
       subtitle = "aggregate by genus",
       caption = "Top 10 taxa")

print(p_top_meta)

#####################################HG########################################
#krakenHG
p_top_krakenHG <- top_krakenHG %>% data.frame() %>% ggplot(aes(fill=OTU, y=av_abundance, x=sample_type)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(title = "KrakenHG",
       subtitle = "aggregate by genus",
       caption = "Top 10 taxa")

print(p_top_krakenHG)

#krakenHG_allknown
p_top_krakenHG_allknown <- top_krakenHG_allknown %>% data.frame() %>% ggplot(aes(fill=OTU, y=av_abundance, x=sample_type)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(title = "KrakenHG_allknown",
       subtitle = "aggregate by genus",
       caption = "Top 10 taxa")

print(p_top_krakenHG_allknown)

#brakenHG
p_top_brakenHG <- top_brakenHG %>% data.frame() %>% ggplot(aes(fill=OTU, y=av_abundance, x=sample_type)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(title = "BrakenHG",
       subtitle = "aggregate by genus",
       caption = "Top 10 taxa")

print(p_top_brakenHG)
```


One table for each phyloseq project with four colums(B1, B4, B5, B7)
```{r}
################################typical#####################################
#kraken
kraken_B1_phy <- PhySeq_Kraken_20.07.21 %>% subset_samples(sample_type=="B1")
kraken_B1_top10_phy <- get_top_taxa(kraken_B1_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
kraken_B1_table <- as.data.frame(tax_table(kraken_B1_top10_phy))

kraken_B4_phy <- PhySeq_Kraken_20.07.21 %>% subset_samples(sample_type=="B4")
kraken_B4_top10_phy <- get_top_taxa(kraken_B4_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
kraken_B4_table <- as.data.frame(tax_table(kraken_B4_top10_phy))

kraken_B5_phy <- PhySeq_Kraken_20.07.21 %>% subset_samples(sample_type=="B5")
kraken_B5_top10_phy <- get_top_taxa(kraken_B5_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
kraken_B5_table <- as.data.frame(tax_table(kraken_B5_top10_phy))

kraken_B7_phy <- PhySeq_Kraken_20.07.21 %>% subset_samples(sample_type=="B7")
kraken_B7_top10_phy <- get_top_taxa(kraken_B7_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
kraken_B7_table <- as.data.frame(tax_table(kraken_B7_top10_phy))

kaken_top_table <- data.frame(kraken_B1_table$genus, kraken_B4_table$genus, kraken_B5_table$genus, kraken_B7_table$genus)

#braken
braken_B1_phy <- PhySeq_Braken_20.07.21 %>% subset_samples(sample_type=="B1")
braken_B1_top10_phy <- get_top_taxa(braken_B1_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
braken_B1_table <- as.data.frame(tax_table(braken_B1_top10_phy))

braken_B4_phy <- PhySeq_Braken_20.07.21 %>% subset_samples(sample_type=="B4")
braken_B4_top10_phy <- get_top_taxa(braken_B4_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
braken_B4_table <- as.data.frame(tax_table(braken_B4_top10_phy))

braken_B5_phy <- PhySeq_Braken_20.07.21 %>% subset_samples(sample_type=="B5")
braken_B5_top10_phy <- get_top_taxa(braken_B5_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
braken_B5_table <- as.data.frame(tax_table(braken_B5_top10_phy))

braken_B7_phy <- PhySeq_Braken_20.07.21 %>% subset_samples(sample_type=="B7")
braken_B7_top10_phy <- get_top_taxa(braken_B7_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
braken_B7_table <- as.data.frame(tax_table(braken_B7_top10_phy))

braken_top_table <- data.frame(braken_B1_table$genus, braken_B4_table$genus, braken_B5_table$genus, braken_B7_table$genus)

#meta
meta_B1_phy <- PhySeq_Metaphlan_20.07.21 %>% subset_samples(sample_type=="B1")
meta_B1_top10_phy <- get_top_taxa(meta_B1_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
meta_B1_table <- as.data.frame(tax_table(meta_B1_top10_phy))

meta_B4_phy <- PhySeq_Metaphlan_20.07.21 %>% subset_samples(sample_type=="B4")
meta_B4_top10_phy <- get_top_taxa(meta_B4_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
meta_B4_table <- as.data.frame(tax_table(meta_B4_top10_phy))

meta_B5_phy <- PhySeq_Metaphlan_20.07.21 %>% subset_samples(sample_type=="B5")
meta_B5_top10_phy <- get_top_taxa(meta_B5_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
meta_B5_table <- as.data.frame(tax_table(meta_B5_top10_phy))

meta_B7_phy <- PhySeq_Metaphlan_20.07.21 %>% subset_samples(sample_type=="B7")
meta_B7_top10_phy <- get_top_taxa(meta_B7_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
meta_B7_table <- as.data.frame(tax_table(meta_B7_top10_phy))

meta_top_table <- data.frame(meta_B1_table$genera, meta_B4_table$genus, meta_B5_table$genus, meta_B7_table$genus)

##################################HG#####################################3
#krakenHG
krakenHG_B1_phy <- PhySeq_KrakenHG_06.09.21 %>% subset_samples(sample_type=="B1")
krakenHG_B1_top10_phy <- get_top_taxa(krakenHG_B1_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
krakenHG_B1_table <- as.data.frame(tax_table(krakenHG_B1_top10_phy))

krakenHG_B4_phy <- PhySeq_KrakenHG_06.09.21 %>% subset_samples(sample_type=="B4")
krakenHG_B4_top10_phy <- get_top_taxa(krakenHG_B4_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
krakenHG_B4_table <- as.data.frame(tax_table(krakenHG_B4_top10_phy))

krakenHG_B5_phy <- PhySeq_KrakenHG_06.09.21 %>% subset_samples(sample_type=="B5")
krakenHG_B5_top10_phy <- get_top_taxa(krakenHG_B5_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
krakenHG_B5_table <- as.data.frame(tax_table(krakenHG_B5_top10_phy))

krakenHG_B7_phy <- PhySeq_KrakenHG_06.09.21 %>% subset_samples(sample_type=="B7")
krakenHG_B7_top10_phy <- get_top_taxa(krakenHG_B7_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
krakenHG_B7_table <- as.data.frame(tax_table(krakenHG_B7_top10_phy))

krakenHG_top_table <- data.frame(krakenHG_B1_table$genus, krakenHG_B4_table$genus, krakenHG_B5_table$genus, krakenHG_B7_table$genus)

#brakenHG
brakenHG_B1_phy <- PhySeq_BrakenHG_06.09.21 %>% subset_samples(sample_type=="B1")
brakenHG_B1_top10_phy <- get_top_taxa(brakenHG_B1_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
brakenHG_B1_table <- as.data.frame(tax_table(brakenHG_B1_top10_phy))

brakenHG_B4_phy <- PhySeq_BrakenHG_06.09.21 %>% subset_samples(sample_type=="B4")
brakenHG_B4_top10_phy <- get_top_taxa(brakenHG_B4_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
brakenHG_B4_table <- as.data.frame(tax_table(brakenHG_B4_top10_phy))

brakenHG_B5_phy <- PhySeq_BrakenHG_06.09.21 %>% subset_samples(sample_type=="B5")
brakenHG_B5_top10_phy <- get_top_taxa(brakenHG_B5_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
brakenHG_B5_table <- as.data.frame(tax_table(brakenHG_B5_top10_phy))

brakenHG_B7_phy <- PhySeq_BrakenHG_06.09.21 %>% subset_samples(sample_type=="B7")
brakenHG_B7_top10_phy <- get_top_taxa(brakenHG_B7_phy, n=10, discard_other=T) %>% microbiome::transform(transform = "compositional")
brakenHG_B7_table <- as.data.frame(tax_table(brakenHG_B7_top10_phy))

brakenHG_top_table <- data.frame(brakenHG_B1_table$genus, brakenHG_B4_table$genus, brakenHG_B5_table$genus, brakenHG_B7_table$genus)
```

