---
title: "Parents"
output: html_notebook
---
This notebook includes all parents variables

variable number 
common: 3 (Family_ID, Sample_type, Sample_ID)
diet: 4
technical: 13
background: father 20 
            mother 44
            partner 3
            family 10
            environment 19
            total 96


```{r warning=FALSE, include=FALSE}
library(phyloseq)
library(tidyverse)
library(dplyr)
library(microbiome)
library(hrbrthemes)
library(knitr)
library(viridis)
library(RVAideMemoire)
library(vegan)
```


#1. Data prepare
Mege diet, background and technical variables into one file
Create the phyloseq project

read in data:
```{r}
diet <- read.csv("WGS_parents_diet_09.11.21.csv")
background_all <- read.csv("WGS_background_all_v1.csv")
background_vocabulary <- read_csv("vocabulary.csv", guess_max = 9000)
tech <- read.csv("WGS_technicalVariables_parents_20.10.21.csv")


braken_sample <- read.csv("../WGS/Braken-HumGut/sample_selection_metadata_06.09.2021.csv")
otu <- read.csv("../WGS/Braken-HumGut/count_table_brakenHG_selection_06.09.2021.csv")
taxa <- read.csv("../WGS/Braken-HumGut/taxonomy_brakenHG_06.09.2021.csv")
```


drop off intant variables from background variables
then merge these tables together with left_joint
```{r}
#braken
braken_sample <- braken_sample %>% filter(Sample_type %in% c("F", "M")) %>% 
                  select(!c("family_ID","sample_type","run_ID", 
                                 "file_pair1", "reads_pair1", "file_pair2", "reads_pair2", "all_reads"))

#background
infant_variables <- background_vocabulary %>% filter(Topic == "Birth") %>% select(Variable) %>% pull()
background <- background_all %>%  select(!any_of(infant_variables))

#tech
tech <- tech %>% select(!c("file_pair1", "reads_pair1", "file_pair2", "reads_pair2", "all_reads"))

#table joint
parent_variables <- left_join(braken_sample, tech) %>% left_join(diet) %>% left_join(background)

write_csv(parent_variables, "parent_variables-2021-11-21.csv")
```


Now we are ready for the phyloseq project
```{r}
sample <- parent_variables %>% tibble::column_to_rownames("sample_ID")


TAX = tax_table(as.matrix(taxa))
OTU = otu_table(as.matrix(otu), taxa_are_rows = TRUE)
SAMPLE = sample_data(as.data.frame(sample))

parent_phy <- phyloseq(OTU, TAX, SAMPLE)

parent_phy <- parent_phy %>% aggregate_taxa("species") %>% 
                aggregate_rare(level = "species", detection = 0.1/100, prevalence = 1/100) %>%
                subset_taxa(!(species=="Unknown")) %>% #drop the unknown group
                microbiome::transform(transform = "compositional")


saveRDS(parent_phy, "parent_phy-2021-11-21.rds")
```


#2. Primary exploration

```{r}
#load in files if start here
parent_phy <- read_rds("parent_phy-2021-11-21.rds")
sample <- read.csv("parent_variables-2021-11-21.csv")

#all meaningful variable names
varnames <- sample %>% select(!c("Project", "SubProject", "sample_ID", "Sample_ID", "Family_ID")) %>% colnames()
```


First, we need a vocabulary

```{r}

vocabulary <- data.frame("Variable"= c(1), "Type"=c(1))


for (variables in colnames(sample)) {
  curr_col <- select(sample, variables)
  curr_ele <- unique(curr_col)
  if ("TRUE"%in% curr_ele | "FALSE"%in% curr_ele ) {
    vocabulary <- rbind(vocabulary, c(variables, "Bool"))
  } else if (length(curr_ele[,1]) > 5) {
    vocabulary <- rbind(vocabulary, c(variables, "Numberic"))
  } else {
    vocabulary <- rbind(vocabulary, c(variables, "categorical"))
  }
}
vocabulary <- vocabulary %>% filter(!Variable==1)
write_csv(vocabulary, "parent_vocabulary-2021-11-21.csv", col_names = T)
```


an overview for the valuable fields
```{r}
vocabulary <- read_csv("vocabulary.csv", guess_max = 9000)
var_bocate <- vocabulary %>% filter(Type=="Bool") %>% select(Variable) %>%pull()

overview <- sample %>% select(Family_ID, any_of(var_bocate))
overview <-  data.frame(lapply(overview, as.character))
overview_long <- overview %>% pivot_longer(!Family_ID, names_to = "question", values_to = "answers")
overview_counts <- overview_long %>% group_by(question, answers) %>% tally()
#overview_counts <- overview_counts %>% filter(answers %in% c("TRUE", "FALSE") | is.na(answers))

p_over <- overview_counts %>% ggplot(aes(x=question, y=n, fill=answers)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme(legend.position = "bottom",
        text=element_text(size=10))
p_over
```


First, try PcoA
```{r}
set.seed(4235421)
ord <- ordinate(parent_phy, "MDS", "bray")

varnames <- sample %>% select(!c("Project", "SubProject", "sample_ID", "Sample_ID", "Family_ID")) %>% colnames()

for (variable in varnames) {
  p_test <- plot_ordination(parent_phy, ord, color = variable) +
                geom_point(size=5)
  print(p_test)
}

```


Next we can try PERMNOVA
```{r}
for (variable in varnames) {
  p_test <- plot_landscape(parent_phy, method = "NMDS", distance = "bray", col = variable, size = 3)
  print(p_test)
}
```
sig_variables
```{r}
 
for (variable in sig_adonis$variable) {
  p_test <- plot_landscape(parent_phy, method = "NMDS", distance = "bray", col = variable, size = 3)
  print(p_test)
}
```



The following codes come from Roosa
Let's try to make it work first

step1. loop through variables and time points and collect the data to a data frame
```{r}
list_adonis <- list()
counter <- 0
list_beta <- list()
sample <- sample %>% column_to_rownames("sample_ID")

for (age in unique(sample$Sample_type)) {
  
  for (variable in colnames(sample)) {
    counter = counter + 1
    
    curr_meta <- parent_phy %>% subset_samples(Sample_type == age) %>%
                          sample_data() %>%
                          as("data.frame")
    samples <- rownames(curr_meta[!is.na(curr_meta[variable]),])
    curr_meta <- curr_meta[rownames(curr_meta) %in% samples,]
    ##These two steps together drop the na rows within the "variable" column
    if (nrow(unique(curr_meta[variable])) > 1) {
      
      curr_asv <- parent_phy %>%
        subset_samples(Sample_type == age) %>%
        otu_table()
      
      curr_asv <- curr_asv[,colnames(curr_asv) %in% samples] ##drop those sampleIDs with na in sample table
      curr_dist <- curr_asv[rowSums(curr_asv) != 0,] %>%
        t() %>%
        vegdist(method = "bray") ##give the distance between two different sample group
                                ##Bray-Crutis distance: 0--identical micro-composition
                                ##                      1--totally different micro-composition
      
      current_ado <- adonis(as.formula(paste("curr_dist ~ ", variable, sep = )), curr_meta) 
      current_res <- as.data.frame(current_ado$aov.tab[variable, 1:6])
      current_res$sample_type <- age
      current_res$variable <- variable
      current_res$fdr <- p.adjust(current_res$`Pr(>F)`, method = "fdr")
      list_adonis[[counter]] <- current_res
        
      ## betadispersion 
      m_beta=betadisper(curr_dist, curr_meta[,variable])
      m_beta
      list_beta[[counter]] <- anova(m_beta)
    } else {
      next
    }
  }
}

all_adonis <- do.call(rbind, list_adonis)
```

The betadispersion table
```{r}
list_beta_df <- data.frame()
for (i in 1:length(list_beta)) {
  if (length(list_beta[[i]]) == 0) {
    list_beta_df[nrow(list_beta_df) + 1,] <- NA
    # list_beta_df$variable[nrow(list_beta_df) + 1] <- as.data.frame(list_adonis[i])["variable"]
    # list_beta_df$sample_type[nrow(list_beta_df) + 1] <- as.data.frame(list_adonis[i])["sample_type"]
    
  } else {
    current_beta <- as.data.frame(list_beta[i])["Groups", 1:5]
    current_beta <- as.data.frame(cbind(current_beta, 
                                        as.data.frame(list_adonis[i])[c("sample_type", "variable")]))
    if (length(list_beta_df) == 0) {
      list_beta_df <- current_beta
    } else {
      list_beta_df[nrow(list_beta_df) + 1,] <- current_beta
    }
  }
}


write_csv(all_adonis, "adonis_parents-2021-11-21.csv")
write_csv(list_beta_df, "betadispersion-parents-2021-11-21.csv")
```


step2. Plot significant variables with non-significant betadispersion

For memory reasons, some variables can be removed
```{r}
rm(asv, taxa, sample, infant, envir, family, father, mother, partner)
gc()
```

If run earlier, read in here:
```{r message=FALSE, warning=FALSE, include=FALSE}
all_adonis <- read_csv("adonis_backgound-2021-11-01.csv")
list_beta_df <- read_csv("betadispersion-2021-11-01.csv")
iphy <- read_rds("Braken_hum&BVs-2021-11-01.rds")
```

prepare
```{r}
# select only significant variables
all_adonis <- na.omit(all_adonis) # remove the rows were p-values were not calculated
sig_adonis <- all_adonis[all_adonis$`Pr(>F)` < 0.05,]

# select only variables with non significant beta dispersion
betadisp <- list_beta_df[!is.na(list_beta_df$variable),]

sig_adonis$betadispersion <- NA
```


loop to add the status of betadispersion in adonis
```{r}
for (variable in unique(sig_adonis$variable)) {
  for (age in unique(sig_adonis$sample_type)) {
    if (nrow(sig_adonis[sig_adonis$variable == variable &
                        sig_adonis$sample_type == age,]) > 0) {
      
      #sign status in sig_adonis according to the Pr..F. value
      beta_status <- betadisp[betadisp$variable == variable &
                              betadisp$sample_type == age,]$Pr..F.
      if (is.na(beta_status)) {
        sig_adonis[sig_adonis$variable == variable &
                     sig_adonis$sample_type == age,]$betadispersion <- NA
      } else if (beta_status < 0.05) {
        sig_adonis[sig_adonis$variable == variable &
                     sig_adonis$sample_type == age,]$betadispersion <- TRUE
      } else {
        sig_adonis[sig_adonis$variable == variable &
                     sig_adonis$sample_type == age,]$betadispersion <- FALSE
      }
    }
  }
}

write_csv(sig_adonis, "sig_adonis-parents-2021-11-21.csv")
```


Step3 Plot
```{r}

# plot dispersions in flagged variables

# plot sig variables heatmap?

plot_list <- list()
counter <- 0

for (variable in unique(sig_adonis$variable)) {
  for (age in unique(sig_adonis$sample_type)) {
    if (nrow(sig_adonis[sig_adonis$variable == variable &
                        sig_adonis$sample_type == age,]) > 0) {
      beta_status <- sig_adonis[sig_adonis$variable == variable &
                               sig_adonis$sample_type == age,]$betadispersion
      if (is.na(beta_status)) {
        break
      }
      if (beta_status) {
        counter <- counter + 1
        curr_meta <- iphy %>%
          subset_samples(sample_type == age) %>%
          sample_data() %>%
          as("data.frame")
        samples <- rownames(curr_meta[!is.na(curr_meta[variable]),])
        curr_meta <- curr_meta[rownames(curr_meta) %in% samples,]
        # leave out variables with over 10 levels
        if (nrow(unique(curr_meta[variable])) > 10) {
          break
        } else {
          
          curr_asv <- iphy %>%
            subset_samples(sample_type == age) %>%
            microbiome::transform(transform = "compositional") %>%
            otu_table()
          
          curr_asv <- curr_asv[,colnames(curr_asv) %in% samples]
          curr_dist <- curr_asv[rowSums(curr_asv) != 0,] %>%
            t() %>%
            vegdist(method = "bray")
          m_beta=betadisper(curr_dist, curr_meta[,variable])
          
          curr_plot <- plot(m_beta, main = paste(age, variable, sep = " "))
          plot_list[[counter]] <- curr_plot
          #print(plot(m_beta, main = paste(age, variable, sep = " ")))
          png(file = paste("plot/", age, variable, ".png", seq = " "))
          plot(m_beta, main = paste(age, variable, sep = " "))
          dev.off()
        }
      }
    }
  }
}
```




Step4 Pairwise comparisons
```{r}
pair_ado <- list()

for (variable in unique(sig_adonis$variable)) {
  for (age in unique(sig_adonis$sample_type)) {
    if (nrow(sig_adonis[sig_adonis$variable == variable &
                        sig_adonis$sample_type == age,]) > 0) {
      
      curr_meta <- iphy %>% subset_samples(sample_type == age) %>%
                            sample_data() %>%
                            as("data.frame")
      
      if (variable %in% colnames(curr_meta)) {
        if (length(unique(curr_meta[,variable])) > 10 |
            length(unique(na.omit(curr_meta[,variable]))) == 2) {
          next
        } else {
          samples <- rownames(curr_meta[!is.na(curr_meta[variable]),])
          curr_meta <- curr_meta[rownames(curr_meta) %in% samples,]
          
          curr_asv <- iphy %>% subset_samples(sample_type == age) %>%
                               microbiome::transform(transform = "compositional") %>%
                               otu_table()
          
          curr_asv <- curr_asv[,colnames(curr_asv) %in% samples]
          curr_dist <- curr_asv[rowSums(curr_asv) != 0,] %>%
            t() %>%
            vegdist(method = "bray")
          result <- pairwise.perm.manova(curr_dist, curr_meta[,variable], test = "Pillai", nperm = 9999, 
                                         progress = TRUE, p.method = "fdr")
          print(result)
          temp_df <- as.data.frame(result$p.value)
          p_df <- expand.grid(as.vector(dimnames(result$p.value)[1][[1]]),
                              as.vector(dimnames(result$p.value)[2][[1]]))
          # Collect all the p-values to the data frame
          p_list <- vector()
          for(col in colnames(temp_df)) {
            p_list <- c(p_list, temp_df[[col]])
          }
          p_df$pval <- p_list
          p_df$variable <- variable
          p_df$timepoint <- age
          # save to list
          pair_ado[[paste(age, variable)]] <- p_df
        }
      }
    }
  }
}

```



Step4 collect all result
```{r}
all_pair_adonis <- do.call(rbind, pair_ado)

write_csv(all_pair_adonis, "pairwise-2021-11-01.csv")
```

