---
title: "Braken_BF"
output: html_notebook
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(phyloseq)
library(tidyverse)
library(microbiome)
library(vegan)
library(RVAideMemoire)
```

Load files
```{r message=FALSE, warning=FALSE, include=FALSE}
b1b4 <- read_csv("WGS_Breastfeeding_B1toB4_011221.csv", guess_max = 9000)
b5 <- read_csv("WGS_Breastfeeding_B5_011221.csv", guess_max = 9000)
b6b7 <- read_csv("WGS_Breastfeeding_B6toB7_011221.csv", guess_max = 9000)

#WGS data
braken_sample <- read.csv("../WGS/Braken-HumGut/sample_selection_metadata_06.09.2021.csv")
otu <- read.csv("../WGS/Braken-HumGut/count_table_brakenHG_selection_06.09.2021.csv")
taxa <- read.csv("../WGS/Braken-HumGut/taxonomy_brakenHG_06.09.2021.csv")
```


#Phyloseq project prepare
Note: the codes below only analysis the infant samples

First the sample table
```{r}
#infant sample
braken_infant_sample <- braken_sample %>% filter(!(sample_type %in% c("F", "M")))


sample <- braken_infant_sample %>% left_join(b1b4, by = "sample_ID") %>% 
                                   left_join(b5, by = "sample_ID") %>%
                                   left_join(b6b7, by = "sample_ID") %>%
                            column_to_rownames("sample_ID")
```

Now the phyloseq project
```{r}
TAX = tax_table(as.matrix(taxa))
OTU = otu_table(as.matrix(otu), taxa_are_rows = TRUE)
SAMPLE = sample_data(as.data.frame(sample))

phy <- phyloseq(OTU, TAX, SAMPLE)

#Before further opperation, aggregate at species level
iphy <- phy %>% aggregate_taxa("species") %>% 
                aggregate_rare(level = "species", detection = 0.1/100, prevalence = 1/100) %>%
                subset_taxa(!(species=="Unknown")) #drop the unknown group

iphy <- iphy %>% microbiome::transform(transform = "compositional")

saveRDS(iphy, "Braken&BF-2021-12-10.rds")
```

 
#PERMANOVA test
The following code originally from rossa

adoins
```{r}
list_adonis <- list()
counter <- 0
list_beta <- list()

for (age in unique(sample$sample_type)) {
  
  for (variable in colnames(sample)) {
    counter = counter + 1
    
    curr_meta <- iphy %>% subset_samples(sample_type == age) %>%
                          sample_data() %>%
                          as("data.frame")
    samples <- rownames(curr_meta[!is.na(curr_meta[variable]),])
    curr_meta <- curr_meta[rownames(curr_meta) %in% samples,]
    ##These two steps together drop the na rows within the "variable" column
    if (nrow(unique(curr_meta[variable])) > 1) {
      
      curr_asv <- iphy %>%
        subset_samples(sample_type == age) %>%
        otu_table()
      
      curr_asv <- curr_asv[,colnames(curr_asv) %in% samples] ##drop those sampleIDs with na in sample table
      curr_dist <- curr_asv[rowSums(curr_asv) != 0,] %>%
        t() %>%
        vegdist(method = "bray") ##give the distance between two different sample group
                                ##Bray-Crutis distance: 0--identical micro-composition
                                ##                      1--totally different micro-composition
      
      current_ado <- adonis(as.formula(paste("curr_dist ~ ", variable, sep = )), curr_meta) 
      current_res <- as.data.frame(current_ado$aov.tab[variable, 1:6])
      current_res$sample_type <- age
      current_res$variable <- variable
      #current_res$fdr <- p.adjust(current_res$`Pr(>F)`, method = "fdr")
      list_adonis[[counter]] <- current_res
        
      ## betadispersion 
      m_beta=betadisper(curr_dist, curr_meta[,variable])
      m_beta
      list_beta[[counter]] <- anova(m_beta)
    } else {
      next
    }
  }
}

all_adonis <- do.call(rbind, list_adonis)
all_adonis$fdr <- p.adjust(all_adonis$`Pr(>F)`, method = "fdr")
```


betadispersion
```{r}
list_beta_df <- data.frame()
for (i in 1:length(list_beta)) {
  if (length(list_beta[[i]]) == 0) {
    list_beta_df[nrow(list_beta_df) + 1,] <- NA
    # list_beta_df$variable[nrow(list_beta_df) + 1] <- as.data.frame(list_adonis[i])["variable"]
    # list_beta_df$sample_type[nrow(list_beta_df) + 1] <- as.data.frame(list_adonis[i])["sample_type"]
    
  } else {
    current_beta <- as.data.frame(list_beta[i])["Groups", 1:5]
    current_beta <- as.data.frame(cbind(current_beta, 
                                        as.data.frame(list_adonis[i])[c("sample_type", "variable")]))
    if (length(list_beta_df) == 0) {
      list_beta_df <- current_beta
    } else {
      list_beta_df[nrow(list_beta_df) + 1,] <- current_beta
    }
  }
}


write_csv(all_adonis, "adonis_BF-2021-12-11.csv")
write_csv(list_beta_df, "betadispersion_BF-2021-12-11.csv")
```


#significant variables plot

remove some variables to save memory
```{r}
gc()
```


read the necessary files again 
```{r message=FALSE, warning=FALSE, include=FALSE}
all_adonis <- read_csv("adonis_BF-2021-12-11.csv")
list_beta_df <- read_csv("betadispersion_BF-2021-12-11.csv")
iphy <- read_rds("Braken&BF-2021-12-10.rds")
```


prepare and the sig_adoin file
```{r}
# select only significant variables
all_adonis <- na.omit(all_adonis) # remove the rows were p-values were not calculated
sig_adonis <- all_adonis[all_adonis$`Pr(>F)` < 0.05,]

# select only variables with non significant beta dispersion
betadisp <- list_beta_df[!is.na(list_beta_df$variable),]

sig_adonis$betadispersion <- NA
```


loop to add the status of betadispersion in adonis
```{r}
for (variable in unique(sig_adonis$variable)) {
  for (age in unique(sig_adonis$sample_type)) {
    if (nrow(sig_adonis[sig_adonis$variable == variable &
                        sig_adonis$sample_type == age,]) > 0) {
      
      #sign status in sig_adonis according to the Pr..F. value
      beta_status <- betadisp[betadisp$variable == variable &
                              betadisp$sample_type == age,]$Pr..F.
      if (is.na(beta_status)) {
        sig_adonis[sig_adonis$variable == variable &
                     sig_adonis$sample_type == age,]$betadispersion <- NA
      } else if (beta_status < 0.05) {
        sig_adonis[sig_adonis$variable == variable &
                     sig_adonis$sample_type == age,]$betadispersion <- TRUE
      } else {
        sig_adonis[sig_adonis$variable == variable &
                     sig_adonis$sample_type == age,]$betadispersion <- FALSE
      }
    }
  }
}

write_csv(sig_adonis, "sig_adonis_BF-2021-12-11.csv")
```


plot loop

(actually, there is no TURE for betadispersal)
```{r}
# plot dispersions in flagged variables

# plot sig variables heatmap?

plot_list <- list()
counter <- 0

for (variable in unique(sig_adonis$variable)) {
  for (age in unique(sig_adonis$sample_type)) {
    if (nrow(sig_adonis[sig_adonis$variable == variable &
                        sig_adonis$sample_type == age,]) > 0) {
      beta_status <- sig_adonis[sig_adonis$variable == variable &
                               sig_adonis$sample_type == age,]$betadispersion
      if (is.na(beta_status)) {
        break
      }
      if (beta_status) {
        counter <- counter + 1
        curr_meta <- iphy %>%
          subset_samples(sample_type == age) %>%
          sample_data() %>%
          as("data.frame")
        samples <- rownames(curr_meta[!is.na(curr_meta[variable]),])
        curr_meta <- curr_meta[rownames(curr_meta) %in% samples,]
        # leave out variables with over 10 levels
        if (nrow(unique(curr_meta[variable])) > 10) {
          break
        } else {
          
          curr_asv <- iphy %>%
            subset_samples(sample_type == age) %>%
            microbiome::transform(transform = "compositional") %>%
            otu_table()
          
          curr_asv <- curr_asv[,colnames(curr_asv) %in% samples]
          curr_dist <- curr_asv[rowSums(curr_asv) != 0,] %>%
            t() %>%
            vegdist(method = "bray")
          m_beta=betadisper(curr_dist, curr_meta[,variable])
          
          curr_plot <- plot(m_beta, main = paste(age, variable, sep = " "))
          plot_list[[counter]] <- curr_plot
          #print(plot(m_beta, main = paste(age, variable, sep = " ")))
          png(file = paste("plot/", age, variable, ".png", seq = " "))
          plot(m_beta, main = paste(age, variable, sep = " "))
          dev.off()
        }
      }
    }
  }
}
```

#Pairwise comparions
```{r}
pair_ado <- list()

for (variable in unique(sig_adonis$variable)) {
  for (age in unique(sig_adonis$sample_type)) {
    if (nrow(sig_adonis[sig_adonis$variable == variable &
                        sig_adonis$sample_type == age,]) > 0) {
      
      curr_meta <- iphy %>% subset_samples(sample_type == age) %>%
                            sample_data() %>%
                            as("data.frame")
      
      if (variable %in% colnames(curr_meta)) {
        if (length(unique(curr_meta[,variable])) > 10 |
            length(unique(na.omit(curr_meta[,variable]))) == 2) {
          next
        } else {
          samples <- rownames(curr_meta[!is.na(curr_meta[variable]),])
          curr_meta <- curr_meta[rownames(curr_meta) %in% samples,]
          
          curr_asv <- iphy %>% subset_samples(sample_type == age) %>%
                               microbiome::transform(transform = "compositional") %>%
                               otu_table()
          
          curr_asv <- curr_asv[,colnames(curr_asv) %in% samples]
          curr_dist <- curr_asv[rowSums(curr_asv) != 0,] %>%
            t() %>%
            vegdist(method = "bray")
          result <- pairwise.perm.manova(curr_dist, curr_meta[,variable], test = "Pillai", nperm = 9999, 
                                         progress = TRUE, p.method = "fdr")
          print(result)
          temp_df <- as.data.frame(result$p.value)
          p_df <- expand.grid(as.vector(dimnames(result$p.value)[1][[1]]),
                              as.vector(dimnames(result$p.value)[2][[1]]))
          # Collect all the p-values to the data frame
          p_list <- vector()
          for(col in colnames(temp_df)) {
            p_list <- c(p_list, temp_df[[col]])
          }
          p_df$pval <- p_list
          p_df$variable <- variable
          p_df$timepoint <- age
          # save to list
          pair_ado[[paste(age, variable)]] <- p_df
        }
      }
    }
  }
}

```


data recording
```{r}
all_pair_adonis <- do.call(rbind, pair_ado)

write_csv(all_pair_adonis, "pairwise_technical-2021-11-09.csv")
```

